---
title: "1.03_Intro.Rmd"
output: html_document
---

```{r setup, include=FALSE}
# Install cran packages
list.of.packages <- c("igraph")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(igraph)
```


# What are soical networks?

Network graph
adjacency matrix
edgelist

igraph r package

library(igraph)
g <- graph.edgelist(as.matrix(df), directed = FALSE)

V(g) gives the vertices for the igraph object and lists them as well

E(g) gives the number of edges for the igraph object and lists them as well 

gorder(g) simply get the number of vertices
gsize(g) get the number of edges

plot(g) to get visualization of the network

```{r}
friends<- read.csv("../Input/friends.csv")
head(friends)

# Convert friends dataframe to a matrix
friends.mat <- as.matrix(friends)

# Convert friends matrix to an igraph object
g <- graph.edgelist(friends.mat, directed = FALSE)

# Subset vertices and edges
V(g)
E(g)

# Count number of edges
gsize(g)

# Count number of vertices
gorder(g)

# Plot
plot(g)
```

## Network attributes:

Network attributes can be categorical or numerical

Vertex and edge attributes

Edges - weights 
The weight may refer to the number of contacts between 2 people

## To add attributes to an igraph object:
```{r}
# Adding Vertex Attributes

g <- set_vertex_attr(
  g, 
  "age", 
  value = c(
    20,25,21,23,24,23,22
    )
  )

vertex_attr(g)

# Adding Edge Attributes

g <- set_edge_attr(
  g, 
  "frequency", 
  value = c(
    2,1,1,1,3,2,4
    )
  )

edge_attr(g)
```

## Alternatively, you may already have the attributes in your input dataframe

```{r}
edges.df <- read.csv("../Input/friends1_edges.csv")
vertices.df <- read.csv("../Input/friends1_nodes.csv")

graph_from_data_frame(d=edges.df, vertices=vertices.df, directed=FALSE)
```


## subsetting networks:

```{r}
E(g)[[inc('Britt')]]

E(g)[[frequency>=3]]
```


## You can also change the visual attributes of networks:

```{r}
# Network visualization
V(g)$color <- ifelse(
  V(g)$age > 22, "red", "white"
)

w1 <- E(g)$hours


# Plot the network varying edges by weights
m1 <- layout_nicely(g)
plot(g, 
        vertex.label.color = "black", 
        edge.color = 'black',
        edge.width = w1,
        layout = m1)
```
