---
title: "2.03_Identifying_important_vertices_in_a_network.Rmd"
output: html_document
---

```{r setup, include=FALSE}
# Install cran packages
list.of.packages <- c("igraph")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(igraph)
```


# Directed Networks

The simplest measure of a vertexes importance is degree

In directed, you have in.degree and out.degree

Test if an edge exists between 2 vertices:
```{r}
# Is there an edge between Britt and Rene?
g['Britt','Rene']

# Show all edges to or from Britt
incident(g, 'Britt', mode=c("all"))

# Find the starting vertex of all edges:
head_of(g,E(g))
```


### Read in the measles graph info 
```{r}
measles <- read.csv("../Input/measles.csv")
g_measles <- graph_from_data_frame(measles, directed = TRUE)

# is the graph directed?
is.directed(g_measles)

# Is the graph weighted?
is.weighted(g_measles)

# Where does each edge originate from?
table(head_of(g_measles, E(g_measles)))
```


### Explore the measles network
```{r}
# Make a basic plot
plot(g_measles, 
     vertex.label.color = "black", 
     edge.color = 'gray77',
     vertex.size = 0,
     edge.arrow.size = 0.1,
     layout = layout_nicely(g_measles))

# Is there an edge going from vertex 184 to vertex 178?
g_measles['184', '178']

# Is there an edge going from vertex 178 to vertex 184?
g_measles['178', '184']
```


### Identifying neighbors:
```{r}
neighbors(g_measles,"184",mode=c("all"))

# Finding intersections (i.e. if 2 vertices don't share a direct neighbor)

x<- neighbors(g_measles, "45",mode=c("all"))

y<- neighbors(g_measles, "183",mode=c("all"))

intersection(x,y)
```


### Paths:

Geodesic distance: how far two nodes are from one another

Diameter - the longest possible path in a network
```{r}
farthest_vertices(g_measles)

get_diameter(g_measles)

# Identifying vertices reachable in N steps
ego(g_measles, 2, "184", mode=c("out"))

# Identify vertices that can reach vertex 184 within two connections
ego(g_measles, 2, '184', mode = c('in'))
```



### Important and influential vertices

Measures of vertex importance
degree
betweenness
eigenvector centrality
closeness centrality
pagerank centrality

## Calculating degree
```{r}
degree(g_measles, mode=c("all"))

# Calculate the out-degree of each vertex
g.outd <- degree(g_measles, mode = c("out"))

# View a summary of out-degree
table(g.outd)

# Make a histogram of out-degrees
hist(g.outd, breaks = 30)

# Find the vertex that has the maximum out-degree
which.max(g.outd)
```

### Betweenness:
This measures how frequently a given vertex serves as the shortest path between any pair of vertices in a network
```{r}
betweenness(g_measles, directed="TRUE")

# Normalize by
betweenness(g_measles, directed=TRUE,normalized=TRUE)

# Calculate betweenness of each vertex
g.b <- betweenness(g_measles, directed = TRUE)

# Show histogram of vertex betweenness
hist(g.b, breaks = 80)

# Create plot with vertex size determined by betweenness score
plot(g, 
     vertex.label = NA,
     edge.color = 'black',
     vertex.size = sqrt(g.b)+1,
     edge.arrow.size = 0.05,
     layout = layout_nicely(g_measles))
```

### Visualizing important nodes and edges
Use make_ego_graph() to create a subset of our network comprised of vertices that are connected to vertex 184. The first argument is the original graph g. The second argument is the maximal number of connections that any vertex needs to be connected to our vertex of interest. In this case we can use diameter() to return the length of the longest path in the network. The third argument is our vertex of interest which should be 184. The final argument is the mode. In this instance you can include all connections regardless of direction.
```{r}
# Make an ego graph
g184 <- make_ego_graph(g_measles, diameter(g_measles), nodes = '184', mode = c("all"))[[1]]

# Get a vector of geodesic distances of all vertices from vertex 184 
dists <- distances(g184, "184")

# Create a color palette of length equal to the maximal geodesic distance plus one.
colors <- c("black", "red", "orange", "blue", "dodgerblue", "cyan")

# Set color attribute to vertices of network g184.
V(g184)$color <- colors[dists+1]

# Visualize the network based on geodesic distance from vertex 184 (patient zero).
plot(g184, 
     vertex.label = dists, 
     vertex.label.color = "white",
     vertex.label.cex = .6,
     edge.color = 'black',
     vertex.size = 7,
     edge.arrow.size = .05,
     main = "Geodesic Distances from Patient Zero"
     )
```

###

```{r}


```





















