---
title: "2.03_Identifying_important_vertices_in_a_network.Rmd"
output: html_document
---

```{r setup, include=FALSE}
# Install cran packages
list.of.packages <- c("igraph")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(igraph)
```

### Characterizing network structures:
eigenvector centrality: a measure of how well a vertex is connected to other vertices
```{r}
# eigenvector centrality 
eigen_centrality(g_measles)$vector

# Edge Density - a proportion of edges that do exist, out of how many that could exist given a number of nodes
edge_density(g_measles)

# Average path length - mean of the lengths of the shortest paths between all pairs of vertices
mean_distance(g_measles, directed=TRUE)
# Lower mean distance means more interconnected
```

### Forrest Gump Network
```{r}
gump<-read.csv("../Input/gump.csv")

# Inspect Forrest Gump Movie dataset
head(gump)

# Make an undirected network
g_gump <- graph_from_data_frame(gump, directed = FALSE)

# Identify key nodes using eigenvector centrality
g.ec <- eigen_centrality(g_gump)
which.max(g.ec$vector)

# Plot Forrest Gump Network
plot(g_gump,
vertex.label.color = "black", 
vertex.label.cex = 0.6,
vertex.size = 25*(g.ec$vector),
edge.color = 'gray88',
main = "Forrest Gump Network"
)
```
### Network density and average path length
```{r}
# Get density of a graph
gd <- edge_density(g_gump)

# Get the diameter of the graph g
diameter(g_gump, directed = FALSE)

# Get the average path length of the graph g
g.apl <- mean_distance(g_gump, directed = FALSE)
g.apl
```

### Understanding network structures

Random network generation is used to compare the metrics of a given network - used to whether features of your original network are particularly unusual.

```{r}
random<-erdos.renyi.game(n=gorder(g_gump),p.or.m=edge_density(g_gump),type="gnp")
plot(g_gump)
edge_density(g_gump)
mean_distance(g_gump, directed = FALSE)

plot(random)
edge_density(random)
#Get the average path length of the random graph g.random
mean_distance(random, directed = FALSE)
```
You can see that the mean distance is shorter in the gump network, suggesting it's more interconnected than a random network

## Random graphs & randomization tests

1. Generate 1000 random graphs based on the original network - e.g. with the same number of vertices and approximate density.
2. Calculate the average path length of the original network.
3. Calculate the average path length of the 1000 random networks.
4. Determine how many random networks have an average path length greater or less than the original network's average path length.

## Generate 1000 random graphs:
```{r}
# Generate 1000 random graphs
gl <- vector('list',1000)

for(i in 1:1000){
    gl[[i]] <- erdos.renyi.game(
                    n = gorder(g_gump), 
                    p.or.m = edge_density(g_gump), 
                    type = "gnp"
                     ) 
}

## Calculate average path length of 1000 random graphs:
gl.apls <- unlist(
  lapply(gl, mean_distance, directed = FALSE) 
)
```

## Plot the graphs
```{r}
## Comparing to the original network
hist(gl.apls, xlim = range(c(1.5, 6)))
# Plot the distribution of average path lengths
abline(v = mean_distance(g_gump, directed=FALSE),col = "red", lty = 3,lwd = 2)
# Calculate the proportion of graphs with an average path length lower than our observed
mean(gl.apls < mean_distance(g_gump, directed=FALSE))
```

## Network substructures:

Transitivity: assess the probability that the adjacent vertices of a given vertex are connected

local transitivty: calculates the proportion of closed triangles a given vertex is a part of, out of the total number of theoretical triangles a vertex could be a part of

Transitivity: for every 3 nodes there a multiple possibilities for how the interconnections look

transitivity() measures the probability that the adjacent vertices are connected

You can also calculate the local transitivity that will calculate the proportion of connections that are closed

Cliques - a structure in which all vertices connect to each other

Typically you don't see a whole network being a clique - instead they typically exist within a network

largest_clique() - identify the largest clique in the network

max_cliques() - identify returns list of cliques that are the biggest - these are also not a part of a larger clique

```{r}
matrix(triangles(g_gump), nrow=3)

# Count the number of triangles that vertex "BUBBA" is in.
count_triangles(g_gump, vids='BUBBA')

# Calculate  the global transitivity of the network.
g.tr <- transitivity(g_gump)
g.tr

# Calculate the local transitivity for vertex BUBBA.
transitivity(g_gump, vids='BUBBA', type = "local")
```

### Transitivity Randomizations:
As you did for the average path length, let's investigate if the global transitivity of the Forrest Gump network is significantly higher than we would expect by chance for random networks of the same size and density. You can compare Forrest Gump's global transitivity to 1000 other random networks.

```{r}
# Calculate average transitivity of 1000 random graphs
gl.tr <- lapply(gl,transitivity)
gl.trs <- unlist(gl.tr)

# Get summary statistics of transitivity scores
summary(gl.trs)

# Calculate the proportion of graphs with a transitivity score higher than Forrest Gump's network
mean(gl.trs > g.tr)
=======
# find all of the closed "triangles of a graph"
triangles(g_gump)
# calculate transitivity:
transitivity(g_gump)
# calculate local transitivity for FORREST
transitivity(g_gump,vids="FORREST",type="local")
```

### Identifying cliques:

Cliques: substructures where all of the vertices are connected 

largest_cliques finds all largest cliques in the input graph. A clique is largest if there is no other clique including more vertices.
```{r}
# Identify the largest clique
largest_cliques(g_gump)
>>>>>>> b2aa1321feee488b265e39a34b3cc7bad9b71a9c
```

### Cliques:
Identifying cliques is a common practice in undirected networks. In a clique every two unique nodes are adjacent - that means that every individual node is connected to every other individual node in the clique. In this exercise you will identify the largest cliques in the Forrest Gump network. You will also identify the number of maximal cliques of various sizes. A clique is maximal if it cannot be extended to a larger clique.

```{r}
# Identify the largest cliques in the network
largest_cliques(g_gump)

# Determine all maximal cliques in the network and assign to object 'clq'
clq <- max_cliques(g_gump)

# Calculate the size of each maximal clique.
table(unlist(lapply(clq, length)))

```

### Visualize largest cliques

Often in network visualization you will need to subset part of a network to inspect the inter-connections of particular vertices. Here, you will create a visualization of the largest cliques in the Forrest Gump network. In the last exercise you determined that there were two cliques of size 9. You will plot these side-by-side after creating two new igraph objects by subsetting out these cliques from the main network. The function subgraph() enables you to choose which vertices to keep in a new network object.

```{r}
# Assign largest cliques output to object 'lc'
lc <- largest_cliques(g_gump)

# Create two new undirected subgraphs, each containing only the vertices of each largest clique.
gs1 <- as.undirected(subgraph(g_gump, lc[[1]]))
gs2 <- as.undirected(subgraph(g_gump, lc[[2]]))


# Plot the two largest cliques side-by-side

par(mfrow=c(1,2)) # To plot two plots side-by-side

plot(gs1,
     vertex.label.color = "black", 
     vertex.label.cex = 0.9,
     vertex.size = 0,
     edge.color = 'gray28',
     main = "Largest Clique 1",
     layout = layout.circle(gs1)
)

plot(gs2,
     vertex.label.color = "black", 
     vertex.label.cex = 0.9,
     vertex.size = 0,
     edge.color = 'gray28',
     main = "Largest Clique 2",
     layout = layout.circle(gs2)
)
```







max_cliques finds all maximal cliques in the input graph. A clique in maximal if it cannot be extended to a larger clique. The largest cliques are always maximal, but a maximal clique is not neccessarily the largest.
```{r}
# Find max size:
max_cliques(g_gump)
```



