---
title: "2.03_Identifying_important_vertices_in_a_network.Rmd"
output: html_document
---

```{r setup, include=FALSE}
# Install cran packages
list.of.packages <- c("igraph")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(igraph)
```

### Characterizing network structures:
eigenvector centrality: a measure of how well a vertex is connected to other vertices
```{r}
# eigenvector centrality 
eigen_centrality(g_measles)$vector

# Edge Density - a proportion of edges that do exist, out of how many that could exist given a number of nodes
edge_density(g_measles)

# Average path length - mean of the lengths of the shortest paths between all pairs of vertices
mean_distance(g_measles, directed=TRUE)
# Lower mean distance means more interconnected
```

### Forrest Gump Network
```{r}
gump<-read.csv("../Input/gump.csv")

# Inspect Forrest Gump Movie dataset
head(gump)

# Make an undirected network
g_gump <- graph_from_data_frame(gump, directed = FALSE)

# Identify key nodes using eigenvector centrality
g.ec <- eigen_centrality(g_gump)
which.max(g.ec$vector)

# Plot Forrest Gump Network
plot(g_gump,
vertex.label.color = "black", 
vertex.label.cex = 0.6,
vertex.size = 25*(g.ec$vector),
edge.color = 'gray88',
main = "Forrest Gump Network"
)
```
### Network density and average path length
```{r}
# Get density of a graph
gd <- edge_density(g_gump)

# Get the diameter of the graph g
diameter(g_gump, directed = FALSE)

# Get the average path length of the graph g
g.apl <- mean_distance(g_gump, directed = FALSE)
g.apl
```

### Understanding network structures

Random network generation is used to compare the metrics of a given network - used to whether features of your original network are particularly unusual.

```{r}
random<-erdos.renyi.game(n=gorder(g_gump),p.or.m=edge_density(g_gump),type="gnp")
plot(g_gump)
edge_density(g_gump)
mean_distance(g_gump, directed = FALSE)

plot(random)
edge_density(random)
#Get the average path length of the random graph g.random
mean_distance(random, directed = FALSE)
```
### you can see that the mean distance is shorter in teh gump network, suggesting it's more interconnected than a random network

Random graphs & randomization tests

1. Generate 1000 random graphs based on the original network - e.g. with the same number of vertices and approximate density.
2. Calculate the average path length of the original network.
3. Calculate the average path length of the 1000 random networks.
4. Determine how many random networks have an average path length greater or less than the original network's average path length.

## Generate 1000 random graphs:
```{r}
# Generate 1000 random graphs
gl <- vector('list',1000)

for(i in 1:1000){
    gl[[i]] <- erdos.renyi.game(
                    n = gorder(g_gump), 
                    p.or.m = edge_density(g_gump), 
                    type = "gnp"
                     ) 
}

## Calculate average path length of 1000 random graphs:
gl.apls <- unlist(
  lapply(gl, mean_distance, directed = FALSE) 
)
```

```{r}
## Comparing to the original network
hist(gl.apls, xlim = range(c(1.5, 6)))
# Plot the distribution of average path lengths
abline(v = mean_distance(g_gump, directed=FALSE),col = "red", lty = 3,lwd = 2)
# Calculate the proportion of graphs with an average path length lower than our observed
mean(gl.apls < mean_distance(g_gump, directed=FALSE))

```

### Network substructures:

```{r}


```




